---
title: "Analysis of Prices"
author: "Floriano Peixoto"
output:
  pdf_document: default
  html_notebook: default
  html_document: default
---

Here we are going to analyse the prices each competitor gives over time in a data set and discover what it has give us.

The data is described as bellow:

- PROD_ID: Product ID. the values varies between P1 to P9;

- DATE_EXTRACTION: Date and Time of the extraction of the competitors’ price, under YYYY-MM-DD HH:MM:SS format;

- COMPETITOR: Competitors’ ID (C1 to C6);

- COMPETITOR_PRICE: Competitors’ price per product, which can depend on the payment method;

- PAY_TYPE: Payment Method (1=deferred payment, 2=immediate payment)

```{r, include=FALSE}
library(lubridate)
library(dplyr)
library(psych)
library(ggplot2)
library(scales)
source("http://peterhaschke.com/Code/multiplot.R")

```

```{r}
prices <-  read.csv("comp_prices.csv", sep=",", stringsAsFactors = T)

tail(prices)
```

As the data documentation tells us the competitor is monitored twice a day.

```{r}
prices <- arrange(prices, COMPETITOR, PROD_ID, DATE_EXTRACTION)

tail(prices)
```

Ordering the set by COMPETITOR, PROD_ID and DATE_EXTRACTION we can see that it´s true, we can turn that into a single line displaying it as min and max price.

```{r}
prices_min_max <- mutate(prices, YEAR = lubridate::year(DATE_EXTRACTION), 
                            MONTH = lubridate::month(DATE_EXTRACTION),
                            DAY = lubridate::day(DATE_EXTRACTION),
                            PROD_ID = as.factor(PROD_ID)) %>%
                  group_by(COMPETITOR, PROD_ID, YEAR, MONTH, DAY) %>%
                  summarise(MIN_PRICE = min(COMPETITOR_PRICE),
                            MAX_PRICE = max(COMPETITOR_PRICE)) %>%
                  arrange(COMPETITOR, PROD_ID, YEAR, MONTH, DAY)

tail(prices_min_max)
```

##Distribuition of products by competitor

With the data in our hands we can try to discrinate the presence of competitors in the market showing the products it sells.

```{r}
presence_market <- group_by(prices_min_max, COMPETITOR, PROD_ID) %>%
                   summarise(M = n())

tail(presence_market)
```

```{r}
g <- ggplot(presence_market, aes(y = M, x = COMPETITOR))

g +  geom_bar(stat = "identity", aes(fill = PROD_ID)) 
```

The plot show us the there are 4 majors competitors C1, C2, C3 and C6, the minor one is C5. It does not seems to have one great competitor, but we have a weaker one in the set. It seems that every major compoetitor sells every product, but the other two has some missing products.


##The competitor with lowest price by product

With our data we can see now what are the competitors with the best price by product.

```{r}
competitor_best_price <- group_by(prices_min_max, PROD_ID, COMPETITOR) %>%
                         summarise(PRICE = min(MIN_PRICE)) %>%
                         slice(which.min(PRICE))

competitor_best_price
```

```{r}
g <- ggplot(competitor_best_price, aes(y = PROD_ID, x = COMPETITOR))

g +  geom_bar(stat = "identity", aes(fill = PROD_ID), position = "dodge") 
```


The competitor C1 has the best prices for almost all products, loosing to C4 in just two products P2 and P4.

##The competitor with the highest prices

Now let´s see the competitor with the highest prices.

```{r}
competitor_worst_price <- group_by(prices_min_max, PROD_ID, COMPETITOR) %>%
                         summarise(PRICE = max(MAX_PRICE)) %>%
                         slice(which.max(PRICE))

competitor_worst_price
```

```{r}
g <- ggplot(competitor_worst_price, aes(y = PROD_ID, x = COMPETITOR))

g +  geom_bar(stat = "identity", aes(fill = PROD_ID), position = "dodge")
```

Diferently from the lowest prices, we can see a better distribuition over the competitors. 

##The lowest and higher prices per competitor by product

It´s intereting to compare the prices each competitor has, historically, for each product in terms to understand more about their behaviors during the year.

```{r}
comp_best_worst_pric_prod <- group_by(prices_min_max, COMPETITOR, PROD_ID) %>%
                             summarise(MIN_PRICE = min(MIN_PRICE),
                                       MAX_PRICE = max(MAX_PRICE)) %>%
                             arrange(COMPETITOR, PROD_ID)

comp_best_worst_pric_prod
```

```{r}

for(product in levels(comp_best_worst_pric_prod$PROD_ID)){
  
  g <- ggplot(filter(comp_best_worst_pric_prod, PROD_ID == product), aes(y = MAX_PRICE, x = COMPETITOR, ymin = MIN_PRICE, ymax = MAX_PRICE, fill = COMPETITOR))
  
  g <- g +  geom_bar(stat = "identity") + 
  geom_errorbar()   + 
  ggtitle(product)

  
  print(g)
}

```

As we can see some competitors have similar prices for some products, but others have a really lower costs. And as we could see before not all of them sell all products, so some of them has higher presence in the market.
